Profit Optimisation in Heterogeneous 2-link Parallel Games

### 2024-08-25 thesis/article.tex
Thesis: Adds Bibliography with Harks and Cole references

%\usepackage{tikz}
\usepackage{enumerate}

@@

We know from [insert ref] that regardless of the distribution function $a$ or the latency functions (as long as they are non-decreasing) a $\NE$ always exists.
In that $\NE$ the users are split with the money-sensitive ones on the link with the lower toll and the time-sensitive ones on the higher toll one, each one seeing their link as either lower or equal cost with the other (while they all see traffic equally).

@@

Consider a 2-link parallel network with latency functions $l_1(x), l_2(x)$ and tolls $t=(t_1, t_2)$ assigned to them.
Without loss of generality we assume $t_1 > t_2$ which in turn means that for the resulting equilibrium flow $x = (x_1(t), x_2(t))$ it holds that $l_1(x_1(t)) < l_2(x_2(t))$.
This equilibrium is guaranteed to exist according to [TODO: add ref]).
For the 2 edges of the game we introduce the term time-money sensitivity split $\as(t)$ as the value for which if any player $p$ has $\alpha(p)=\as(t)$ then that player sees the $2$ edge total latencies as equal.
More specifically, $\as(t)$ is the value of the distribution function $\alpha$ for which it holds
\[l_1(x_1(t)) + \as(t) t_1 = l_2(x_2(t)) + \as(t) t_2\]
Solving for $\as$ we get
\[\as(t) = \frac{l_2(x_2(t)) - l_1(x_1(t))}{t_1 - t_2}\]
\\
Before we get into a more formal definition and description of $\as$, it'd help to first discuss the nature of the split that $\as$ captures.
Regardless of whether there exists a player $p$ such that $\alpha(p) = \as(t)$, it still holds that
\begin{itemize}
	\item if $\alpha(p) < \as$ then $l_1(x_1(t)) + \alpha(p) t_1 < l_2(x_2(t)) + \alpha(p) t_2$, thus $p$ is on edge $1$
	\item if $\alpha(p) > \as$ then $l_1(x_1(t)) + \alpha(p) t_1 > l_2(x_2(t)) + \alpha(p) t_2$, thus $p$ is on edge $2$
	\item if $\alpha(p) = \as$ then $l_1(x_1(t)) + \alpha(p) t_1 = l_2(x_2(t)) + \alpha(p) t_2$, thus $p$ is either on edge $1$ or $2$
\end{itemize}
Remember that we have assumed that $t_1 > t_2 \Rightarrow l_1(x_1(t)) < l_2(x_2(t))$.
Therefore the flow passing through the faster edge $1$ is contained with players where $\alpha(p) \le \as(t)$, while the flow in the cheaper edge $2$ is contained with players where $\alpha(p) \ge \as(t)$.
Since the players are sorted in increasing order according to $\alpha$, it helps to view the players in the first case as the "lower" part of the split and respectively the players in the second case as the "higher" part of the split.
[TODO: Alternative splits: rushed-relaxed, rich-poor]
Table \autoref{table:split_summary} provides of a summary for each part of the split

@@

Finally to acknowledge that we have arbitrarily handled any players with $\alpha(p) = \as(t)$.
Those players, if any, even though they see both edges with equal total cost, in the optimal flow $x$ they have settled in either edge $1$ or $2$.
We investigate more in depth the relation of $\as$ to the flow in [TODO: add ref].
\\[12pt]
We can now formally define $\as$ and prove some of its properties.
\begin{definition}
	Given a 2-link parallel heterogeneous game with latency functions $l_1, l_2$, we define the time-money sensitivity split function $\as: \R^2_+ \rightarrow (0, +\infty)$ where for given tolls $t = (t_1, t_2)$
	\[\as(t) = \frac{l_2(x_2(t)) - l_1(x_1(t))}{t_1 - t_2}\]
\end{definition}
The function is well defined since the given tolls $t$ result in a single optimal flow $x(t) = (x_1(t), x_2(t))$ [TODO: add ref].

@@

In this section we will examine how different distribution functions $\alpha$ can affect the profit maximisation game between the toll owners.
[TODO: add framework for examining games]
We will restrict our analysis on distribution functions such that $\alpha(0) > 0$, the reason being that we want to avoid two instances.
The first is to ensure that $\alpha(p) > 0$ when $p > 0$, while the second is that $0$ is not an accumulation point.
[TODO: maybe mention add ref who also take the first restriction as it creates games with no optimal tolls]
The following two examples demonstrate the issues that might occur if these restrictions are not met.

@@

\chapter*{Bibliography}
%\begin{thebibliography}{99}
%\end{thebibliography}
TODO

>>>>

%\usepackage{tikz}
\usepackage{enumerate}
\usepackage{natbib}

@@

We know from Cole et al. \cite{10.1145/780542.780618} that regardless of the distribution function $a$ or the latency functions (as long as they are non-decreasing) a $\NE$ always exists.
In that $\NE$ the users are split with the money-sensitive ones on the link with the lower toll and the time-sensitive ones on the higher toll one, each one seeing their link as either lower or equal cost with the other (while they all see traffic equally).

@@

Consider a 2-link parallel network with latency functions $l_1(x), l_2(x)$ and tolls $t=(t_1, t_2)$ assigned to them.
Without loss of generality we assume $t_1 > t_2$ which in turn means that for the resulting equilibrium flow $x = (x_1(t), x_2(t))$ it holds that $l_1(x_1(t)) < l_2(x_2(t))$.
This equilibrium is guaranteed to exist according to Harks et al. \cite{Harks_2019}.
For the 2 edges of the game we introduce the term time-money sensitivity split $\as(t)$ as the value for which if any player $p$ has $\alpha(p)=\as(t)$ then that player sees the $2$ edge total latencies as equal.
More specifically, $\as(t)$ is the value of the distribution function $\alpha$ for which it holds
\[l_1(x_1(t)) + \as(t) t_1 = l_2(x_2(t)) + \as(t) t_2\]
Solving for $\as$ we get
\[\as(t) = \frac{l_2(x_2(t)) - l_1(x_1(t))}{t_1 - t_2}\]
\\
Before we get into a more formal definition and description of $\as$, it'd help to first discuss the nature of the split that $\as$ captures.
Regardless of whether there exists a player $p$ such that $\alpha(p) = \as(t)$, it still holds that
\begin{itemize}
	\item if $\alpha(p) < \as$ then $l_1(x_1(t)) + \alpha(p) t_1 < l_2(x_2(t)) + \alpha(p) t_2$, thus $p$ is on edge $1$
	\item if $\alpha(p) > \as$ then $l_1(x_1(t)) + \alpha(p) t_1 > l_2(x_2(t)) + \alpha(p) t_2$, thus $p$ is on edge $2$
	\item if $\alpha(p) = \as$ then $l_1(x_1(t)) + \alpha(p) t_1 = l_2(x_2(t)) + \alpha(p) t_2$, thus $p$ is either on edge $1$ or $2$
\end{itemize}
Remember that we have assumed that $t_1 > t_2 \Rightarrow l_1(x_1(t)) < l_2(x_2(t))$.
Therefore the flow passing through the faster edge $1$ is contained with players where $\alpha(p) \le \as(t)$, while the flow in the cheaper edge $2$ is contained with players where $\alpha(p) \ge \as(t)$.
Since the players are sorted in increasing order according to $\alpha$, it helps to view the players in the first case as the "lower" part of the split and respectively the players in the second case as the "higher" part of the split.
[TODO: Alternative splits: rushed-relaxed, rich-poor]
Table \autoref{table:split_summary} provides of a summary for each part of the split

@@

Finally to acknowledge that we have arbitrarily handled any players with $\alpha(p) = \as(t)$.
Those players, if any, even though they see both edges with equal total cost, in the optimal flow $x$ they have settled in either edge $1$ or $2$.
We investigate more in depth the relation of $\as$ to the flow in [TODO: add ref].
\\[12pt]
We can now formally define $\as$ and prove some of its properties.
\begin{definition}
	Given a 2-link parallel heterogeneous game with latency functions $l_1, l_2$, we define the time-money sensitivity split function $\as: \R^2_+ \rightarrow (0, +\infty)$ where for given tolls $t = (t_1, t_2)$
	\[\as(t) = \frac{l_2(x_2(t)) - l_1(x_1(t))}{t_1 - t_2}\]
\end{definition}
The function is well defined since the given tolls $t$ result in a single optimal flow $x(t) = (x_1(t), x_2(t))$ \cite{Harks_2019}.

@@

In this section we will examine how different distribution functions $\alpha$ can affect the profit maximisation game between the toll owners.
[TODO: add framework for examining games]
We will restrict our analysis on distribution functions such that $\alpha(0) > 0$, the reason being that we want to avoid two instances.
The first is to ensure that $\alpha(p) > 0$ when $p > 0$, while the second is that $0$ is not an accumulation point.
[TODO: maybe mention \cite{10.1145/780542.780618} where they also take the first restriction as it creates games with no optimal tolls]
The following two examples demonstrate the issues that might occur if these restrictions are not met.

@@

\bibliographystyle{plain}
\bibliography{ref.bib}
